\documentclass{article}

\usepackage{amsmath}
\usepackage[none]{hyphenat}

\begin{document}
\title{Report: Simulation of a Vibrating System}
\author{Sean Prasertsit}
\date{2014 November 22}
\maketitle

\section{Equations satisfied by the system using Newton's second law of motion.}
The equations are the result of the object and its mass experiencing forces from two horizontal sides. Spring values affect the distance travelled from both sides.
\subsection{}	\(m_1\ddot{x}_1 = -k_1x_1 + k_2(x_2 - x_1)\)
\subsection{}	\(m_2\ddot{x}_2 = -k_2(x_2 - x_1) + k_3(x_3 - x_2)\)
\subsection{}	\(m_3\ddot{x}_3 = -k_3(x_3 - x_2) - k_4x_3\)

\section{System of equations expressed as an eigen problem.}
\(k\) and \(m\) are matrices representing spring values and mass respectively. Thus, \newline
\(
	\begin{bmatrix}
	\frac{k}{m}
	\end{bmatrix}
	\begin{bmatrix}
	x_1 \\
	x_2 \\
	x_3
	\end{bmatrix}
	=
	\begin{bmatrix}
	\ddot{x}_1 \\
	\ddot{x}_2 \\
	\ddot{x}_3
	\end{bmatrix}\). \newline
Without damping, we choose a purely oscillatory solution. \newline
\(x = ve^{i\omega t} \longrightarrow \ddot{x} = -\omega^2ve^{i\omega t} = -\omega^2x \longrightarrow Ax = -\omega^2x\) where \(A\) is the spring and mass matrix. This becomes an eigen problem where \(\lambda = -\omega^2\) so
\(Av = \lambda v\). \newline \newline
\(A = 
	\begin{bmatrix}
	\frac{-k_1 - k_2}{m_1} 	& \frac{k_2}{m_1} 		& 0 \\
	\frac{k_2}{m_2} 		& \frac{-k_2 - k_3}{m_2} & \frac{k_3}{m_2} \\
	0						& \frac{k_3}{m_3}		& \frac{-k_3 - k_4}{m_3}
	\end{bmatrix} \)
where \textit{v} is an eigenvector and \(\lambda \) is a corresponding scalar eigenvalue.

\section{Eigenvector matrix and eigenvalue matrix.}
\(v = 
	\begin{bmatrix}
	 0.1163	&	0.8321	&	0.3803 \\
	 -0.8495	&	-0.2305	&	0.6768 \\
	 0.5146	&	-0.5045	&	0.6303
	\end{bmatrix} \)
\(\lambda =
	\begin{bmatrix}
	-25.7567	&	0		&	0 \\
	0		&	-8.1925	&	0 \\
	0		&	0		&	-3.0508
	\end{bmatrix} \)

\section{Frequencies and amplitudes.}
Frequency is the value \(\omega = \sqrt{-\lambda}\). It is the number of occurrences in one period. Amplitude is an eigenvector, \(X\) of the form \(AX = \lambda X\) and is the measure of change over a period.
\(\omega = 
	\begin{bmatrix}
	4.7121	\\	
	2.7994	\\
	1.8145
	\end{bmatrix} \)
\end{document}